Simple repression draft

### 2020-05-21 doc/section_03_beyond_means.tex
Added to constit prom Bayesian inference + minor file reorg
Update section_03_beyond_means.tex

% !TEX root = ./busty_transcription.tex
\section{Beyond means}
\subsection{A one-state promoter with bursting is the ``best'' model of constitutive promoters}
Before we can tackle simple repression, we need an adequate phenomenological
model of constitutive expression. The literature abounds with options from which
we can choose, and we show several potential kinetic models for constitutive
promoters in~\fig{fig:constit_cartoons}. Let us consider the suitability of each
model for our purposes in turn.

@@

\subsubsection{Model 1 - Poisson promoter.}
This is the picture from Jones et.\ al.~\cite{Jones2014}, which assumes that
transcripts are produced as a Poisson process from a single promoter state. This
model insists that the ``true'' mRNA distribution is Poisson, implying the Fano
factor $\nu$ must be 1. In~\cite{Jones2014}, the authors carefully attribute
deviations from Fano = 1 to intensity variability in fluorescent spots, gene
copy number variation, and copy number fluctuations of the transcription
machinery, e.g., RNAP itself. In this picture, the master equation makes no
appearance, and all the corrections to Poisson behavior are derived as additive
corrections to the Fano factor. For disproving the ``universal noise curve''
from So et al~\cite{So2011}, this picture was excellent. It is appealing in its
simplicity, with only two parameters, the initiation rate $r$ and degradation
rate $\gamma$. Since $\gamma$ is independently known from other experiments, and
the mean mRNA copy number is $r/\gamma$, $r$ is easily inferred from data. In
other words, the model is not excessively complex for the data at hand. But for
many interesting questions, for instance in the recent work~\mmnote{cite
Manuel's preprint once posted}, knowledge of means and variances alone is
insufficient, and a description of the full distribution of molecular counts is
necessary. For this we need a( slightly) more complex model than model 1 that
would allow us to incorporate the non-Poissonian features of constitutive
promoters directly into a master equation formulation.

>>>>

% !TEX root = ./busty_transcription.tex
\section{Beyond Means in Gene Expression}

In this section, our objective is to explore the same models considered in
the previous section, but now with reference to the the question of
how well they describe the distribution of gene expression levels, with
special reference to the variance in these distributions.  To that end,
we repeat the same pattern as in the previous section by examining
the models one by one.

\subsection{A one-state promoter with bursting is the ``best'' model of constitutive promoters}
Before we can tackle simple repression, we need an adequate phenomenological
model of constitutive expression. The literature abounds with options from which
we can choose, and we show several potential kinetic models for constitutive
promoters in~\fig{fig:constit_cartoons}. Let us consider the suitability of each
model for our purposes in turn.

@@

\subsubsection{Noise in the Poisson Promoter Model}
The simplest model of constitutive expression that we can imagine
is shown as model 1 in Figure~\ref{fig:constit_cartoons} and
assumes that
transcripts are produced as a Poisson process from a single promoter state. 
This is the picture from Jones et.\ al.~\cite{Jones2014} that was used to
interpret a systematic study of gene expression noise over a series of
promoters designed to have different strengths.   This
model insists that the ``true'' mRNA distribution is Poisson, implying the Fano
factor $\nu$ must be 1. In~\cite{Jones2014}, the authors carefully attribute
deviations from Fano = 1 to intensity variability in fluorescent spots, gene
copy number variation, and copy number fluctuations of the transcription
machinery, e.g., RNAP itself. In this picture, the master equation makes no
appearance, and all the corrections to Poisson behavior are derived as additive
corrections to the Fano factor. For disproving the ``universal noise curve''
from So et al~\cite{So2011}, this picture was excellent. It is appealing in its
simplicity, with only two parameters, the initiation rate $r$ and degradation
rate $\gamma$. Since $\gamma$ is independently known from other experiments, and
the mean mRNA copy number is $r/\gamma$, $r$ is easily inferred from data. In
other words, the model is not excessively complex for the data at hand. But for
many interesting questions, for instance in the recent work~\mmnote{cite
Manuel's preprint once posted}, knowledge of means and variances alone is
insufficient, and a description of the full distribution of molecular counts is
necessary. For this we need a (slightly) more complex model than model 1 that
would allow us to incorporate the non-Poissonian features of constitutive
promoters directly into a master equation formulation.
